\begin{abstract}
If autoresearch is itself a form of research, then autoresearch can be applied
to research \emph{itself}.
We take this idea literally:
we use an autoresearch loop to optimize the autoresearch loop.

Every existing autoresearch system---from Karpathy's single-track loop to
AutoResearchClaw's multi-batch extension and EvoScientist's persistent
memory---was improved by a \emph{human} who read the code, identified a
bottleneck, and wrote new code.
We ask whether an LLM can do the same, autonomously.

We present \textbf{\methodname{}}, a bilevel framework where an outer loop
meta-optimizes the inner autoresearch loop by generating and injecting new
search mechanisms as Python code at runtime.
The inner loop optimizes the task; the outer loop optimizes how the inner
loop searches.
Both loops use the same LLM---no stronger model is needed at the meta level.

On Karpathy's GPT pretraining benchmark, the meta-autoresearch outer loop
achieves a 5$\times$ improvement over the standard inner loop alone
($-0.045$ vs.\ $-0.009$ \valbpb{}), while parameter-level adjustment without
mechanism change yields no reliable gain.
The outer loop autonomously discovers mechanisms from combinatorial
optimization, multi-armed bandits, and design of experiments---without human
specification of which domains to explore.
These mechanisms succeed by breaking the inner loop's deterministic search
patterns, forcing exploration of directions the LLM's priors systematically avoid.

The core principle is simple:
\emph{autoresearch can meta-autoresearch itself.}
\end{abstract}
