\section{Discussion}
\label{sec:discussion}

\subsection{Hypothesis Testing}

The experimental design is motivated by three hypotheses.

\textbf{H1 (Group~B $>$ Group~A): Not supported.}
Group~B's mean improvement ($-0.006 \pm 0.006$) is numerically worse than
Group~A's ($-0.009 \pm 0.002$), though the difference is not meaningful given
$n=3$.
The outer loop (Level~1.5) increases search diversity---Group~B explores more
parameters than Group~A---but this diversity does not translate into larger
improvements within the 30-iteration budget.
Group~B's R1 achieved essentially zero improvement ($-0.000$), the worst
outcome of any repeat in any group.
The outer loop correctly froze stalled parameters but, having done so, the LLM
found nothing better in the remaining search space.

\textbf{H2 (Group~C $>$ Group~B): Supported.}
Group~C's mean absolute improvement ($-0.045 \pm 0.030$) is 7.5$\times$
Group~B's ($-0.006 \pm 0.006$).
Despite high variance ($\pm 0.030$), two of three repeats produced dramatic
improvements ($-0.065$, $-0.058$), and the separation between Groups~C and~B
is large relative to the within-group variance, providing meaningful evidence
that Level~2 adds value beyond Level~1.5.

\textbf{H3 (Level~2 discovers novel mechanisms autonomously): Supported.}
Across three independent repeats, Level~2 generated mechanisms from three
distinct active domains (combinatorial optimization, online learning, DOE)
without being told which domains to consider (a fourth domain, Bayesian
optimization, was attempted but reverted due to a missing dependency).
Code generation succeeded on the first attempt in all six sessions (zero
retries).
Five of six mechanisms passed import validation and were activated.

\subsection{Why Level~1.5 Alone Is Insufficient}

Level~1.5 can diagnose search problems---it correctly identifies when a
parameter has been attempted repeatedly without improvement---and it can
redirect search by adjusting which parameters are active.
However, it is structurally limited to operating within the existing search
framework.
It cannot change how proposals are generated, modify the acceptance criterion,
add pre-filtering steps, or introduce new search paradigms such as tabu lists,
bandit selection, or orthogonal sampling.
Only Level~2 can produce such code changes.
This structural ceiling explains why H1 is not supported: parameter
redirection alone is insufficient to escape the LLM's deterministic
proposal patterns.

\subsection{Why Group~C's R2 Underperformed}

Group~C's R2 achieved only $-0.011$ improvement, comparable to Groups~A and~B,
despite receiving Level~2 mechanisms.
The most likely explanation is mechanism quality: R2's Level~2 generated the
Multi-Scale Bandit Proposer, which---while valid and correctly injected---may
be less effective than R1's Tabu Search Manager or R3's Orthogonal Exploration
at forcing exploration of the batch size dimension.
A secondary factor is overhead: each Level~2 research session requires
approximately 3~minutes of wall time (four LLM calls), reducing effective inner
iterations.
With two sessions per repeat, Group~C has roughly 6~minutes less search time
than Groups~A and~B, a minor but non-zero cost.

\subsection{Limitations}

\paragraph{Small sample size.}
Three repeats per group is insufficient for rigorous statistical comparison.
Group~C's standard deviation ($\pm 0.030$) is 67\% of its absolute mean, indicating
high variability.
Reliable estimates would require $n \geq 10$ repeats per group.

\paragraph{Baseline variance.}
Baseline \valbpb{} varies across repeats (1.094--1.114) due to training
randomness from data ordering and weight initialization.
Using $\Delta = \text{best} - \text{baseline}$ mitigates this but does not
eliminate it; a lower baseline gives less headroom for improvement.
Future work should use fixed random seeds or report baseline-normalized metrics.

\paragraph{Single benchmark.}
All results are on one task: GPT pretraining at 50M parameters with a 300-second
budget on RTX~5090.
Generalization to other model sizes, training budgets, or tasks is unproven.

\paragraph{Dynamic load fragility.}
A preliminary run was invalidated because the Level~2 dynamic loading pipeline
contained a \texttt{sys.modules} registration bug, causing all mechanism
injections to silently fall back to the original runner.
We fixed the bug before conducting the three reported repeats.
This episode highlights the fragility of runtime code injection: silent
fallback without error is a dangerous failure mode.

\paragraph{External dependency exposure.}
Level~2 has no constraint preventing it from importing external libraries.
One of six generated mechanisms (GP Regressor) required \texttt{sklearn}, which
was not installed.
The validate-and-revert mechanism handled this correctly, but the exposure to
arbitrary dependencies remains a reliability risk.

\subsection{Unexpected Findings}

\paragraph{Group~A's search is near-deterministic.}
Given the same baseline \texttt{train.py}, DeepSeek generates almost the same
proposal sequence across all three repeats.
Group~A's three repeats are therefore not independent samples in the usual
sense: they are three runs of the same near-deterministic process, perturbed
only by training noise.
This determinism is itself a finding: LLMs hold strong priors about which
hyperparameter changes to try first, and the inner loop alone cannot escape
those priors.

\paragraph{Level~2 selects diverse domains without guidance.}
The Level~2 prompt instructs the model to ``propose mechanism improvements''
but does not suggest which algorithmic domains to consider.
Across three independent repeats, the model independently chose combinatorial
optimization, online learning, and design of experiments (plus a Bayesian
optimization attempt that was reverted due to a missing dependency).
The LLM's breadth of knowledge therefore enables genuine exploration at the
mechanism level, not just at the configuration level.

\subsection{Future Work}

Key directions for extending this work include:
(1) scaling to $n \geq 10$ repeats with fixed random seeds for statistical
    power;
(2) pre-installing common scientific libraries (e.g., \texttt{sklearn}) on experiment
    servers to expand Level~2's mechanism vocabulary;
(3) evaluating on multiple benchmarks (different model sizes, tasks, budgets)
    to assess generalization;
(4) adding a fourth group (Level~1 with \texttt{simple\_mode=False}) to
    isolate the contribution of pre-baked mechanisms vs.\ autonomously
    discovered ones; and
(5) investigating whether Level~2's code generation quality can be improved
    by providing a richer interface specification or a test harness.
