\section{Introduction}
\label{sec:introduction}

Large language models have demonstrated a striking capacity for self-directed
scientific iteration: given a task, an LLM can propose a change, execute an
experiment, observe the outcome, and decide whether to keep or discard the
change.
When repeated, this propose--execute--evaluate loop constitutes a form of
automated research~\citep{karpathy2026autoresearch}.
Instantiated for neural network hyperparameter search, we call this loop
\emph{autoresearch}.

Despite its promise, autoresearch as currently practiced has a fundamental
limitation: the search \emph{mechanism} is fixed at design time.
Every system in the literature uses a human-engineered architecture.
\citet{karpathy2026autoresearch} introduced the single-track inner loop with a
keep/discard acceptance rule.
AutoResearchClaw~\citep{aiminglab2026claw} extended it with multi-batch parallel
search.
EvoScientist~\citep{evoscientist2026} added persistent experience memory across
runs.
A human designed each improvement by reading the prior system's code,
identifying a bottleneck, and writing new code to address it.
The systems themselves cannot perform this operation.

This raises a natural question: \emph{can an outer loop perform that same
design step autonomously?}
Can the system read its own inner loop's code, identify structural
bottlenecks, generate new Python code to address them, and inject that
code at runtime---all without human involvement?

We answer this question affirmatively.
We present \textbf{\methodname{}}, a bilevel framework with three operational
levels:
\begin{itemize}
  \item \textbf{Level~1} (inner loop) runs the standard autoresearch cycle: an
        LLM proposes a configuration change, the model trains for a fixed budget,
        validation loss (\valbpb{}) is measured, and the change is kept or
        discarded.
  \item \textbf{Level~1.5} (outer loop) executes every five inner iterations.
        It analyzes the accumulated search trace, freezes parameters that have
        failed repeatedly, unfreezes under-explored parameters, and injects
        strategic guidance text into the next inner cycle.
  \item \textbf{Level~2} (mechanism research) executes every two outer cycles.
        It conducts a structured 4-round research session---Explore, Critique,
        Specify, Generate---producing Python code that implements a novel search
        mechanism.
        The code is patched into the inner loop's runner module via
        \texttt{importlib} dynamic loading and validated before activation.
\end{itemize}

Crucially, all three levels use the same LLM (DeepSeek \texttt{deepseek-chat}).
Any improvement in search effectiveness therefore comes from the bilevel
\emph{architecture}, not from access to a more capable model.

We evaluate four experimental groups on Karpathy's GPT pretraining benchmark:
Group~A (Level~1 only), Group~B (Levels~1 and~1.5), Group~C (all three
levels), and Group~D (Levels~1 and~2, without Level~1.5).
The primary metric is \valbpb{} improvement within a 30-iteration budget using a
300-second training window per iteration, matching the original benchmark
specification.
All groups start from the same baseline configuration and use identical GPU
hardware (RTX~5090).

\paragraph{Contributions.}
\begin{enumerate}
  \item We formalize autoresearch as a bilevel optimization problem in which the
        upper level optimizes the search \emph{mechanism} (runner code) and the
        lower level optimizes the task (training configuration), and demonstrate
        that the upper level can be solved by LLM code generation.
  \item We implement Level~2 mechanism research via a 4-round LLM dialogue and
        runtime code injection, establishing a concrete realization of a
        self-modifying autoresearch system.
  \item We report a controlled ablation showing that Level~2 produces a
        5$\times$ improvement over the Level-1 baseline ($-0.045 \pm 0.030$
        vs.\ $-0.009 \pm 0.002$), with Level~2 autonomously generating
        mechanisms drawn from four distinct algorithmic domains.
  \item We identify why Level~2 succeeds: the generated mechanisms (Tabu
        Search, Multi-Scale Bandit, Orthogonal Exploration) forced exploration
        of \texttt{TOTAL\_BATCH\_SIZE} \emph{reduction}, a direction that the
        LLM's default search path systematically avoids due to an implicit bias
        toward larger batches.
\end{enumerate}

\Cref{sec:related} reviews related work.
\Cref{sec:methods} describes the framework in detail.
\Cref{sec:results} presents ablation results.
\Cref{sec:discussion} interprets hypotheses, failure modes, and limitations.
\Cref{sec:conclusion} concludes.
